---
title: "PMLproject"
author: "Solen Sanschagrin"
date: "February 28, 2016"
output: html_document
---
#Overview
This report is an example of constructing a prediction model using tree models. It uses data from  [http://groupware.les.inf.puc-rio.br/har] to predict the form of an exercise using information from personal activity devices.

#Data

```{r cache=TRUE, echo=FALSE, results='hide', warning=FALSE}
require(ggplot2)
require(lattice)
require(caret)
require(randomForest)
```

Loading data and splitting the training set into a train and test set for cross validation.
```{r, warning=FALSE, results='hide', message=FALSE}
require(caret)
training<-read.csv("pml-training.csv")
testing<-read.csv("pml-testing.csv")
set.seed(111)
inTrain<-createDataPartition(training[,1],p=0.6,list=FALSE) #divide the training set 
train<-training[inTrain,]                   #by random subsampling
test<-training[-inTrain,]
```

 Elimination of features that are mostly NA values and the first 7 columns of data since the time stamps can be used to match directly to the particular movement made.
```{r, warning=FALSE,results='hide'}
missing<-is.na(training)
unusedFeats<-which(colSums(missing)>0) #cols with NAs have 19216 missing values (data has 19622 rows)
train<-train[,-c(1:7,unusedFeats)]
test<-test[,-c(1:7,unusedFeats)]
testing<-testing[,-c(1:7,unusedFeats)]
```

Hardware limitations currently keep from using the full extent of the `train()` function from the caret package so an example is made here using the three first features (after the time or window references), in this case: `names(train)[1:3]`.

```{r, cache=TRUE,warning=FALSE,results='hide'}

modelfitBelt<-train(classe~.,data=train[,c(1:3,ncol(train))],method="rpart",preProcess="pca")
predictTrain<-predict(modelfitBelt,newdata=train[,1:3])
predictTest<-predict(modelfitBelt,newdata=test[,1:3])
trainSuccess<-sum(predictTrain==train[,86])/length(predictTrain)
testSuccess<-sum(predictTest==test[,86])/length(predictTest)
```
The generated model:
```{r}
modelfitBelt
```
The data was preprocessed using the PCA method from the caret package for a success rate of `r round(trainSuccess,2)` on the `train` subset and of `r round(testSuccess,2)` on the `test` subset. The out of sample error is then estimated as 1-`r round(testSuccess,2)` =`r 1-round(testSuccess,2)`.

